\sysfull

### 2017-11-05 paper/intro.tex
text for virtualization

\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks. Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures. The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent. A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware. A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict. To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency). At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks. At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer. This calls for at-runtime task division of any transiently-powered application.

\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task. This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}. Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively. All tasks have no WAR dependency and are consistent. Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).

\textbf{Trade-off with Virtual Memory Manipulation:} WAR dependencies cannot be complied at runtime for merged (virtual) tasks. Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency. Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}. Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}). How to connect this trade-off with task virtualization remains an open question.

>>>>

\textbf{\replace{Static Task-based Model and its Inefficiency:}{The Need for Adaptive Task Division:}} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks. Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures. The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent. A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware. A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict. To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency). At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks. At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer. This calls for at-runtime \add{adaptive} task division of any transiently-powered application\add{---namely \emph{task virtualization}}.

\begin{figure}
	\includegraphics[width=0.6\columnwidth]{presentation/images/alpaca-war.pdf}
	\caption{\label{fig:virtualization}\color{blue} Adaptive task division at run-time might lead to WAR dependencies that do not exist with static task division.}
\end{figure}

\textbf{\replace{Data Inconsistency of Virtual Task:}{The Challange Behind Task Virtualization:}} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task. This however \replace{induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}}{might introduce new \emph{write-after-read (WAR)} dependencies on the variables in non-volatile memory at \emph{run-time} that can break the atomicity of the \emph{virtual task}. Consider the example depicted in Fig.~\ref{fig:virtualization}: Tasks 1--3 being executed consecutively.} \remove{Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.} All tasks \replace{have no WAR dependency and are consistent}{in this figure are atomic since they do not have WAR dependency on the persistent variables they are accessing, e.g. \emph{x} is only read and \emph{y} is only written within Task 1.} Now, suppose three tasks have been virtualized into one \add{at run-time}, namely Task 4: \replace{ \{y=x+1; z=y+1; x=z+1;\} which has inherent WAR dependency (variable \{\tt{x}\}). }{since now \emph{x} is first read and then written, a WAR dependency on \emph{x} is introduced dynamically at run-time---unfortunately Task 4 is \emph{no more atomic} since its re-execution will not always produce the same results.}

\textbf{\replace{Trade-off with Virtual Memory Manipulation:}{Guaranteeing Atomicity of Virtual Tasks:}} \replace{WAR dependencies cannot be complied at runtime for merged (virtual) tasks.}{Even though WAR dependencies introduced by the static task division can be eliminated at compile time, e.g.~\cite{alpaca}, those introduced at run-time dynamically due to task virtualization cannot.} \replace{Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency.}{Therefore, we require an \emph{execution model} that keeps each virtual task atomic by committing the modified persistent variables at the boundary defined dynamically at run-time---keeping the non-volatile memory unmodified upon a power-interrupt and preserving its consistency.}

\TODO{Below is totally disconnected from the flow}
{\color{red}Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}. Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}). How to connect this trade-off with task virtualization remains an open question.}
